### 2016-11-09 datamining/summery.tex
finished graphical models

\documentclass[12pt,landscape]{extarticle}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}
\usepackage{graphicx}
\newcommand{\indep}{\rotatebox[origin=c]{90}{$\vDash$}}

@@

% This sets page margins to .5 inch if using letter paper, and to 1cm
% if using A4 paper. (This probably isn't strictly necessary.)
% If using another size paper, use default 1cm margins.
\ifthenelse{\lengthtest { \paperwidth = 11in}}
    { \geometry{top=.5in,left=.5in,right=.5in,bottom=.5in} }
    {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
        {\geometry{top=0.5cm,left=0.5cm,right=0.5cm,bottom=0.5cm} }
    }

@@

\section{Baysian networks}
\section{Freq item set mining}
\end{multicols}
\end{document}

>>>>

\documentclass[10pt,landscape]{extarticle}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}
\usepackage{graphicx}
\newcommand{\indep}{\rotatebox[origin=c]{90}{$\vDash$}}

@@

% This sets page margins to .5 inch if using letter paper, and to 1cm
% if using A4 paper. (This probably isn't strictly necessary.)
% If using another size paper, use default 1cm margins.
\ifthenelse{\lengthtest { \paperwidth = 11in}}
    { \geometry{top=.5in,left=.5in,right=.5in,bottom=.5in} }
    {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
        {\geometry{top=0.5cm,left=0.2cm,right=0.2cm,bottom=0.5cm} }
        {\geometry{top=0.5cm,left=0.2cm,right=0.2cm,bottom=0.5cm} }
    }

@@

\subsubsection{Iterative proportional fitting}
Is a converging algorithm
Start with a uniform table. Then set the row margins to the fitted counts
and distribute them proportionally. Then do the same for the columns, keep
doing this until the values converged.

\subsection{Decomosable graphical models}
Decomposable models are graphical models that have explicit formulas for the
maximum likelihood estimates.
This is a convenient property from a computational viewpoint.
If we only have to fit one model this is perhaps not so important,
but when we have little prior knowledge we typically have to search a
potentially large space of possible models.
Decomposable models are very easy to characterize by their independence graphs.
They have triangulated independence graphs: their independence graphs have no
chordless cycles of length greater than three. A cycle is called chordless
if no other than successive pairs of vertices in the cycle are adjacent. 
\subsection{Deviance and likelhood ratio test}
The deviance of a fitted model compares the log-likelihood of the fitted model
to the log-likelihood of the saturated model.
The larger the model deviance, the poorer the fit.
The likelihood of a model $M$ is
$L(\overset{\wedge}{P}(x);n(x))=\underset{x}{\prod}\overset{\wedge}{P}(x)^{n(x)}$
and the log likelihood of a model $M$ is
$\underset{x}{\sum} n(x)\log \overset{\wedge}{P}(x)$
$L^i=L(\overset{\wedge}{P}^{M_i})$ is the value of the log likelihood function
valutated at $\overset{\wedge}{P}^{M_i}$; the ML estimates of $P$ under $M_i$.
Let $M_0 \subseteq M_1$; ie $M_0$ can be obtained from $M_1$ by setting
additional $u$-terms to zero (more restrictions). The deviance between $M_0$ and
$M_1$ is $dev(M_0) - dev(M_1) = 2(L^1 - L^0)$.
For large $N$ $2(L^1-L^0) \approx M_0 \chi^2_v$ where the degrees of freedom $v$
is equal to the number of additional restrictions of $M_0$. 
We reject the null hypothesis that $M_0$ is the true model when:
$2(L^1-L^0)>\chi^2_{v;\alpha}$. 
\section{Baysian networks}

\section{Freq item set mining}
\end{multicols}
\end{document}
